---
title: "Capture recapture analyses of greens for Agnese Mancini"
output: html_notebook
---

This document describes data analysis of Agnese Mancini's data of green turtle capture recapture events of hawksbill turtles using the augmented method (Kery and Schaub). 


Initialize the workspace
```{r}
rm(list=ls())
library(jagsUI)
library(tidyverse)
library(lubridate)
library(reshape)
library(bayesplot)
library(ggridges)

source("Mancini_functions.R")

MCMC.params <- list(n.samples = 50000,
                    n.burnin = 30000,
                    n.thin = 5,
                    n.chains = 5)

```

Bring in the data file and get ready for CJS

```{r}

dat.1 <- get.data("data/GTC_20190725_Tomo_v2.csv")

dat.1 %>% filter(species == "Cm") -> dat.1.Cm

CJS.data <- dat2CJS(dat.1.Cm, save.file = FALSE)

CJS.data$data %>% rownames_to_column(var = "ID") -> data.CJS

# Number of years between occasions. 
cap.dates <- paste0(colnames(CJS.data$data), "-01")
delta.dates <- signif(as.numeric(as.Date(cap.dates[2:length(cap.dates)]) -
                                   as.Date(cap.dates[1:(length(cap.dates)-1)]))/365, 1)

# capture history
CH <- as.matrix(CJS.data$data)

nInd <- nrow(CJS.data$data)

ns <- colSums(CJS.data$data)

# find the first capture date
get.first <- function(x) min(which(x != 0))
f <- apply(CH, 1, get.first)

jags.data <- list(y = CH, 
                  f = f, 
                  nind = nInd, 
                  n.occasions = dim(CH)[2],
                  n = ns, 
                  T = ncol(CJS.data$data),
                  dt = c(0, delta.dates))


# initial values - when removing z, no input
# initsFunction <- function(ch){
#   mean.phi <- rbeta(1,2,2)
#   mean.p <- rbeta(1,2,2)
#   z <- known.state.cjs(ch)
#   b <- rnorm(1, 0, 50)
#   A <- list(mean.phi = mean.phi, mean.p = mean.p, z = z, beta = b)
#   return(A)
# }
```


Then run JAGS:
Pdot model
```{r}

if (!file.exists(paste0("RData/CJS_pdot_Cm_JAGS.rds"))){
  ## parameters to monitor - when this is changed, make sure to change
  ## summary statistics indices at the end of this script. 
  parameters <- c("mean.p", "mean.phi", "N", "deviance")
  
  # inits <- lapply(replicate(MCMC.params$n.chains, 
  #                           CH, 
  #                           simplify = FALSE), 
  #                 initsFunction)
  MCMC.params$model.file = "models/Model_CJS_pdot.txt"
  
  jm.pdot <- jags(data = jags.data,
                  #inits = inits,
                  parameters.to.save= parameters,
                  model.file = MCMC.params$model.file,
                  n.chains = MCMC.params$n.chains,
                  n.burnin = MCMC.params$n.burnin,
                  n.thin = MCMC.params$n.thin,
                  n.iter = MCMC.params$n.samples,
                  DIC = T, 
                  parallel=T)
  
  out.list <- list(jags.out = jm.pdot,
                   jags.data = jags.data)
  
  saveRDS(out.list, file = paste0("RData/CJS_pdot_Cm_JAGS.rds"))
  
} else {
  phidot.pdot <- readRDS(file = "RData/CJS_pdot_Cm_JAGS.rds")
}
```

pt
```{r}
if (!file.exists(paste0("RData/CJS_pt_Cm_JAGS.rds"))){
  MCMC.params$model.file <- "models/Model_CJS_pt.txt"
  
  ## parameters to monitor - when this is changed, make sure to change
  ## summary statistics indices at the end of this script. 
  parameters <- c("p", "mean.phi", "N", "deviance")
  
  jm.pt <- jags(data = jags.data,
                #inits = inits,
                parameters.to.save= parameters,
                model.file = MCMC.params$model.file,
                n.chains = MCMC.params$n.chains,
                n.burnin = MCMC.params$n.burnin,
                n.thin = MCMC.params$n.thin,
                n.iter = MCMC.params$n.samples,
                DIC = T, 
                parallel=T)
  
  out.list <- list(jags.out = jm.pt,
                   jags.data = jags.data)
  
  saveRDS(out.list, 
          file = paste0("RData/CJS_pt_Cm_JAGS.rds"))
  
} else {
  phidot.pt <- readRDS(file = "RData/CJS_pt_Cm_JAGS.rds")

}

```


Try the trap response model:
```{r}

if (!file.exists(paste0("RData/CJS_pTrapResponse_Cm_JAGS.rds"))){
  MCMC.params$model.file <- "models/Model_CJS_pTrapResponse.txt"
  
  m <- CH[, 1:(ncol(CH) - 1)]
  u <- which(m == 0)
  m[u] <- 2
  jags.data$m <- m
  
  ## parameters to monitor - when this is changed, make sure to change
  ## summary statistics indices at the end of this script. 
  parameters <- c("beta", "mean.phi", "deviance")
  
  jm.pt.TrapResponse <- jags(data = jags.data,
                             #inits = inits,
                             parameters.to.save= parameters,
                             model.file = MCMC.params$model.file,
                             n.chains = MCMC.params$n.chains,
                             n.burnin = MCMC.params$n.burnin,
                             n.thin = MCMC.params$n.thin,
                             n.iter = MCMC.params$n.samples,
                             DIC = T, 
                             parallel=T)
  
  out.list <- list(jags.out = jm.pt.TrapResponse,
                   jags.data = jags.data)
  
  saveRDS(out.list, 
          file = paste0("RData/CJS_pTrapResponse_Cm_JAGS.rds"))
  
} else {
  phidot.pTrap <- readRDS(file = "RData/CJS_pTrapresponse_Cm_JAGS.rds")

}

```

Time-since-marking - looking at the transients effect on survival 
```{r}

if (!file.exists(paste0("RData/CJS_PhiTSM_pt_Cm_JAGS.rds"))){
  MCMC.params$model.file <- "models/Model_CJS_PhiTSM_pt.txt"
  
  first.cap <- jags.data$f
  m <- matrix(data = 2, nrow = nrow(CH), ncol = ncol(CH))
  for (k in 1:nrow(m)){
    m[k, first.cap[k]] <- 1
  }
  
  jags.data$m <- m
  
  ## parameters to monitor - when this is changed, make sure to change
  ## summary statistics indices at the end of this script. 
  parameters <- c("gamma", "p", "N", "prop.trans", "deviance")
  
  jm.PhiTSM.pt <- jags(data = jags.data,
                       #inits = inits,
                       parameters.to.save= parameters,
                       model.file = MCMC.params$model.file,
                       n.chains = MCMC.params$n.chains,
                       n.burnin = MCMC.params$n.burnin,
                       n.thin = MCMC.params$n.thin,
                       n.iter = MCMC.params$n.samples,
                       DIC = T, 
                       parallel=T)
  
  out.list <- list(jags.out = jm.PhiTSM.pt,
                   jags.data = jags.data)
  
  saveRDS(out.list, 
          file = paste0("RData/CJS_PhiTSM_pt_Cm_JAGS.rds"))
} else {
  phiTSM.pt <- readRDS(file = "RData/CJS_PhiTSM_pt_Cm_JAGS.rds")

}
```

Time since marking for both phi and p:

```{r}

if (!file.exists(paste0("RData/CJS_PhiTSM_pTSM_Cm_JAGS.rds"))){
  MCMC.params$model.file <- "models/Model_CJS_PhiTSM_pTSM.txt"
  
  first.cap <- jags.data$f
  m <- matrix(data = 2, nrow = nrow(CH), ncol = ncol(CH))
  for (k in 1:nrow(m)){
    m[k, first.cap[k]] <- 1
  }
  
  jags.data$m <- m
  
  ## parameters to monitor - when this is changed, make sure to change
  ## summary statistics indices at the end of this script. 
  parameters <- c("gamma", "beta", "prop.trans", "deviance")
  
  jm.PhiTSM.pTSM <- jags(data = jags.data,
                       #inits = inits,
                       parameters.to.save= parameters,
                       model.file = MCMC.params$model.file,
                       n.chains = MCMC.params$n.chains,
                       n.burnin = MCMC.params$n.burnin,
                       n.thin = MCMC.params$n.thin,
                       n.iter = MCMC.params$n.samples,
                       DIC = T, 
                       parallel=T)
  
  out.list <- list(jags.out = jm.PhiTSM.pTSM,
                   jags.data = jags.data)
  
  saveRDS(out.list, 
          file = paste0("RData/CJS_PhiTSM_pTSM_Cm_JAGS.rds"))
} else {
  phiTSM.pTSM <- readRDS(file = "RData/CJS_PhiTSM_pTSM_Cm_JAGS.rds")

}
```

Time since marking for phi and effort for p:

```{r}

if (!file.exists(paste0("RData/CJS_PhiTSM_pEff_Cm_JAGS.rds"))){
  MCMC.params$model.file <- "models/Model_CJS_PhiTSM_pEffort.txt"
  
  # need to count how many capture events occured per season
  dat.1.Cm %>% select(season, "DATE") %>% 
  group_by(season) %>% #-> tmp3
  summarise(effort = n_distinct(DATE)) -> effort.season
  
  first.cap <- jags.data$f
  m <- matrix(data = 2, nrow = nrow(CH), ncol = ncol(CH))
  for (k in 1:nrow(m)){
    m[k, first.cap[k]] <- 1
  }
  
  jags.data$m <- m
  jags.data$x <- effort.season$effort
  
  ## parameters to monitor - when this is changed, make sure to change
  ## summary statistics indices at the end of this script. 
  parameters <- c("gamma", "p", "mu", 
                  "beta", "N", "prop.trans", 
                  "sigma",  "deviance")
  
  jm.PhiTSM.pEff <- jags(data = jags.data,
                         #inits = inits,
                         parameters.to.save= parameters,
                         model.file = MCMC.params$model.file,
                         n.chains = MCMC.params$n.chains,
                         n.burnin = MCMC.params$n.burnin,
                         n.thin = MCMC.params$n.thin,
                         n.iter = MCMC.params$n.samples,
                         DIC = T, 
                         parallel=T)
  
  out.list <- list(jags.out = jm.PhiTSM.pEff,
                   jags.data = jags.data)
  
  saveRDS(out.list, 
          file = paste0("RData/CJS_PhiTSM_pEff_Cm_JAGS.rds"))
} else {
  phiTSM.pEff <- readRDS(file = "RData/CJS_PhiTSM_pEff_Cm_JAGS.rds")

}
```



pt with minimumn CCL as a covariate for survival. Need to redo the data.

```{r}

MCMC.params <- list(n.samples = 80000,
                    n.burnin = 50000,
                    n.thin = 5,
                    n.chains = 5)

dat.1 <- get.data("data/GTC_20190725_Tomo_v2.csv")

dat.1 %>% filter(species == "Cm") -> dat.1.Cm

# Use CCL as a covariate as only three were missed (these may be recaptures also).
dat.1.Cm %>% select(ID, CCL) %>% group_by(ID) %>%
  summarise(min_CCL = min(CCL, na.rm = T)) %>%
  filter(!is.infinite(min_CCL)) -> data.CCL

CJS.data <- dat2CJS(dat.1.Cm, save.file = FALSE)

CJS.data$data %>% rownames_to_column(var = "ID") -> data.CJS

# This gets you just 0s and 1s
data.CCL %>% left_join(data.CJS, by = "ID") %>%
  select(-c("ID", "min_CCL")) -> CH.2

data.CCL  %>% left_join(data.CJS, by = "ID") %>%
  select(min_CCL) -> cov.CCL

# Number of years between occasions. 
cap.dates <- paste0(colnames(CJS.data$data), "-01")
delta.dates <- signif(as.numeric(as.Date(cap.dates[2:length(cap.dates)]) -
                                   as.Date(cap.dates[1:(length(cap.dates)-1)]))/365, 1)

nInd <- nrow(CH.2)
ns <- colSums(CH.2)

# find the first capture date
get.first <- function(x) min(which(x != 0))
f <- apply(CH.2, 1, get.first)

jags.data <- list(y = CH.2, 
                  f = f, 
                  nind = nInd, 
                  n.occasions = dim(CH)[2],
                  n = ns, 
                  T = ncol(CJS.data$data),
                  dt = c(0, delta.dates),
                  x = as.vector(t(cov.CCL)))


if (!file.exists(paste0("RData/CJS_PhiMinCCL_pdot_Cm_JAGS.rds"))){
  MCMC.params$model.file <- "models/Model_CJS_PhiMinCCL_pdot.txt"
  
  ## parameters to monitor - when this is changed, make sure to change
  ## summary statistics indices at the end of this script. 
  parameters <- c("mean.p", "mu", "beta", "sigma", "N", "deviance")
  
  jags.data$x <- 
  jm.PhiMinCCL.pdot <- jags(data = jags.data,
                         #inits = inits,
                         parameters.to.save= parameters,
                         model.file = MCMC.params$model.file,
                         n.chains = MCMC.params$n.chains,
                         n.burnin = MCMC.params$n.burnin,
                         n.thin = MCMC.params$n.thin,
                         n.iter = MCMC.params$n.samples,
                         DIC = T, 
                         parallel=T)
  
  out.list <- list(jags.out = jm.PhiMinCCL.pdot,
                   jags.data = jags.data)
  
  saveRDS(out.list, 
          file = paste0("RData/CJS_PhiMinCCL_pdot_Cm_JAGS.rds"))
  
} else {
  phiMinCCL_pdot <- readRDS(file = "RData/CJS_PhiMinCCL_pdot_Cm_JAGS.rds")
}
```

Trace plots look not so good.  

How about time-varying CCL individual covariate?

```{r}

# if (!file.exists(paste0("RData/CJS_PhiCCL_pt_Cm_JAGS.rds"))){
#   MCMC.params$model.file <- "models/Model_CJS_PhiCCL_pt.txt"
#   dat.1.CCL <- dat2CJS_covCCL(dat.1.Cm)
#   data.CCL %>% left_join(dat.1.CCL, by = "ID") %>%
#     select(-min_CCL) -> dat.2.CCL
# 
# 
#   CCL.mat <- as.matrix(dat.2.CCL[, 2:ncol(dat.2.CCL)])
# 
#   jags.data <- list(y = CH.2,
#                     f = f,
#                     nind = nInd,
#                     n.occasions = dim(CH)[2],
#                     n = ns,
#                     T = ncol(CJS.data$data),
#                     dt = c(0, delta.dates),
#                     x = CCL.mat)
# 
#   ## parameters to monitor - when this is changed, make sure to change
#   ## summary statistics indices at the end of this script.
#   parameters <- c("p", "mu", "beta", "sigma", "N", "deviance")
# 
#   jm.PhiCCL.pt <- jags(data = jags.data,
#                          #inits = inits,
#                          parameters.to.save= parameters,
#                          model.file = MCMC.params$model.file,
#                          n.chains = MCMC.params$n.chains,
#                          n.burnin = MCMC.params$n.burnin,
#                          n.thin = MCMC.params$n.thin,
#                          n.iter = MCMC.params$n.samples,
#                          DIC = T,
#                          parallel=T)
# 
#   out.list <- list(jags.out = jm.PhiCCL.pt,
#                    jags.data = jags.data)
# 
#   saveRDS(out.list,
#           file = paste0("RData/CJS_PhiCCL_pt_Cm_JAGS.rds"))
# }

```


Age class for survival rate:

```{r}

if (!file.exists(paste0("RData/CJS_PhiAgeClass_pt_Cm_JAGS.rds"))){
  MCMC.params$model.file <- "models/Model_CJS_PhiAgeClass_pt.txt"
  dat.1.CCL <- dat2CJS_covCCL(dat.1.Cm)
  data.CCL %>% left_join(dat.1.CCL, by = "ID") %>%
    select(-min_CCL) -> dat.2.CCL

  CCL.mat <- as.matrix(dat.2.CCL[, 2:ncol(dat.2.CCL)])
  
  ageclass.mat <- matrix(1, nrow = nrow(CCL.mat), ncol = ncol(CCL.mat))
  ageclass.mat[which(CCL.mat >= 90.0)] <- 2
  k <- 1
  for ( k in 1:nrow(ageclass.mat)){
    idx.2 <- which(ageclass.mat[k,] == 2)
    if (length(idx.2) > 0){
      ageclass.mat[idx.2[1]:ncol(ageclass.mat)] <- 2
    }     
  }
  
  jags.data <- list(y = CH.2, 
                    f = f, 
                    nind = nInd, 
                    n.occasions = dim(CH)[2],
                    n = ns, 
                    T = ncol(CJS.data$data),
                    dt = c(0, delta.dates),
                    m = ageclass.mat)
  
  ## parameters to monitor - when this is changed, make sure to change
  ## summary statistics indices at the end of this script. 
  parameters <- c("p", "mu", "gamma", "sigma", "N", "deviance")
  
  jm.PhiAgeClass.pt <- jags(data = jags.data,
                         #inits = inits,
                         parameters.to.save= parameters,
                         model.file = MCMC.params$model.file,
                         n.chains = MCMC.params$n.chains,
                         n.burnin = MCMC.params$n.burnin,
                         n.thin = MCMC.params$n.thin,
                         n.iter = MCMC.params$n.samples,
                         DIC = T, 
                         parallel=T)
  
  out.list <- list(jags.out = jm.PhiAgeClass.pt,
                   jags.data = jags.data)
  
  saveRDS(out.list, 
          file = paste0("RData/CJS_PhiAgeClass_pt_Cm_JAGS.rds"))
} else {
  phiAgeClass.pt <- readRDS(file = "RData/CJS_PhiAgeClass_pt_Cm_JAGS.rds")
}

```


```{r}

if (!file.exists(paste0("RData/CJS_PhiAgeClass_TSM_pt_Cm_JAGS.rds"))){
  MCMC.params$model.file <- "models/Model_CJS_PhiAgeClass_TSM_pt.txt"
  dat.1.CCL <- dat2CJS_covCCL(dat.1.Cm)
  data.CCL %>% left_join(dat.1.CCL, by = "ID") %>%
    select(-min_CCL) -> dat.2.CCL

  CCL.mat <- as.matrix(dat.2.CCL[, 2:ncol(dat.2.CCL)])

  CCL.mat[which(CCL.mat >= 90.0)] <- 2
  CCL.mat[which(CCL.mat < 90.0)] <- 1

  first.cap <- jags.data$f
  m <- matrix(data = 2, nrow = nrow(CH), ncol = ncol(CH))
  X <- matrix(data = 1, nrow = nrow(CCL.mat), ncol = ncol(CCL.mat))

  for (k in 1:nrow(m)){
    m[k, first.cap[k]] <- 1
 
  }
  X[which(m == 1 & CCL.mat == 2)] <- 2
  X[which(m == 2 & CCL.mat == 1)] <- 3
  X[which(m == 2 & CCL.mat == 2)] <- 4

  jags.data <- list(y = CH.2,
                    f = f,
                    nind = nInd,
                    n.occasions = dim(CH)[2],
                    n = ns,
                    T = ncol(CJS.data$data),
                    dt = c(0, delta.dates),
                    m = X)

  ## parameters to monitor - when this is changed, make sure to change
  ## summary statistics indices at the end of this script.
  parameters <- c("p", "mu", "gamma", "sigma", "N", "deviance")

  jm.PhiAgeClass.TSM.pt <- jags(data = jags.data,
                                #inits = inits,
                                parameters.to.save= parameters,
                                model.file = MCMC.params$model.file,
                                n.chains = MCMC.params$n.chains,
                                n.burnin = MCMC.params$n.burnin,
                                n.thin = MCMC.params$n.thin,
                                n.iter = MCMC.params$n.samples,
                                DIC = T,
                                parallel=T)

  out.list <- list(jags.out = jm.PhiAgeClass.TSM.pt,
                   jags.data = jags.data)

  saveRDS(out.list,
          file = paste0("RData/CJS_PhiAgeClass_TSM_pt_Cm_JAGS.rds"))
} else {
  phiAgeClass.TSM.pt <- readRDS("RData/CJS_PhiAgeClass_TSM_pt_Cm_JAGS.rds")
}
```
Look at the results and compare which model may be better:

```{r}
phidot.pdot <- readRDS(file = "RData/CJS_pdot_Cm_JAGS.rds")
phidot.pt <- readRDS(file = "RData/CJS_pt_Cm_JAGS.rds")
phidot.pTrap <- readRDS(file = "RData/CJS_pTrapresponse_Cm_JAGS.rds")
phiTSM.pt <- readRDS(file = "RData/CJS_PhiTSM_pt_Cm_JAGS.rds")
phiTSM.pEff <- readRDS(file = "RData/CJS_PhiTSM_pEff_Cm_JAGS.rds")
phiTSM.pTSM <- readRDS(file = "RData/CJS_PhiTSM_pTSM_Cm_JAGS.rds")

# The following didn't do so well wrt trace plots (I think...)
# phiMinCCL.pdot <- readRDS(file = "RData/CJS_PhiMinCCL_pdot_Cm_JAGS.rds")
# phiCCL.pt <- readRDS(file = "RData/CJS_PhiCCL_pt_Cm_JAGS.rds")
# phiAgeClass.pt <- readRDS(file = "RData/CJS_PhiAgeClass_pt_Cm_JAGS.rds")
# phiAgeClass.TSM.pt <- readRDS(file = "RData/CJS_PhiAgeClass_TSM_pt_Cm_JAGS.rds")

```

Look at DICs - although they are not great for model comparison...

```{r}
DICs <- data.frame(model = c("phidot.pdot",  "phidot.pt", "phidot.pTrap", "phiTSM.pt",
                             "phiTSM.pEff", "phiTSM.pTSM"),
                   DIC = c(phidot.pdot$jags.out$DIC, 
                           phidot.pt$jags.out$DIC, 
                           phidot.pTrap$jags.out$DIC,
                           phiTSM.pt$jags.out$DIC, 
                           phiTSM.pEff$jags.out$DIC, 
                           phiTSM.pTSM$jags.out$DIC)) 
                           #phiMinCCL.pdot$jags.out$DIC,
                           #phiAgeClass.pt$jags.out$DIC))
                           #phiAgeClass.TSM.pt$jags.out$DIC))

arrange(DICs, DIC)
```

According to DIC, phi(AgeClass) p(t) is the best.  Let's take a look at the results.

gamma[1] is < 90 cm and gamma[2] is >= 90 cm. Interesting to see that gamma[2] has a lot wider posterior. Median is lower for big ones also... 

```{r}
mcmc_dens(phiAgeClass.pt$jags.out$samples, c("gamma[1]", "gamma[2]"))
```

```{r}
gamma.stats <- phiAgeClass.pt$jags.out$summary[grep("gamma", 
                                               dimnames(phiAgeClass.pt$jags.out$summary)[[1]]),]
gamma.stats
```


```{r}
p.stats <- phiAgeClass.pt$jags.out$summary[grep("p[", 
                                           dimnames(phiAgeClass.pt$jags.out$summary)[[1]],
                                           fixed = TRUE),]
p.stats
```


Survival rates of residents were about 0.6/yr.  Kinda low... 

Using the H-T estimator, we also computed the abundance for each sampling occasion:

```{r}
N.stats <- phiAgeClass.pt$jags.out$summary[grep("N[", 
                                           dimnames(phiAgeClass.pt$jags.out$summary)[[1]],
                                           fixed = TRUE),]

N.stats.df <- data.frame(N.stats)

N.stats.df$season <- colnames(CH.2)

N.stats.df
```


```{r}
ggplot(data = N.stats.df) +
  geom_point(aes(x = season, y = X50.)) + 
  geom_errorbar(aes(x = season, ymin = X2.5., ymax = X97.5.)) + 
  theme(axis.text.x = element_text(angle = 90, hjust = 0.5, vjust = 0.5)) + 
  ylab("Abundance") + xlab("")

```

