---
title: "R Notebook"
output: html_notebook
---

Visually inspecting capture recapture data of green turtles from Agnes Mancini.

```{r}
rm(list=ls())
library(tidyverse)
library(lubridate)
library(reshape)
library(ggridges)

source("Mancini_functions.R")

# from https://stackoverflow.com/questions/5831794/opposite-of-in-exclude-rows-with-values-specified-in-a-vector
`%ni%` <- Negate(`%in%`)

save.fig <- F
```

Pull out one site at a time, then look at how turtles were recaptured. 

```{r}

dat.1.Cm <- get.data.Cm.2023("data/240925_Base datos_Cm_2001-2023 TE.csv")

# Group by community and create CJS data for each community
site.names <- levels(dat.1.Cm$Macro_site)
p.recap <- vector(mode = "numeric")

k <- 1
for (k in 1:length(site.names)){
  dat.1.Cm.site <- dat.1.Cm %>% filter(Macro_site == site.names[k])
  CJS.data <- dat2CJS(dat.1.Cm.site, 
                        save.file = FALSE)
  n.GT2.cap <- length(which(rowSums(CJS.data$data) > 1))
  p.recap[k] <- n.GT2.cap/nrow(CJS.data$data)
    
}

useful.site <- site.names[p.recap > 0.05]

dat.1.Cm %>% filter(Macro_site %in% useful.site) -> dat.2.Cm
loc.stats <- list()
for (k in 1:length(useful.site)){
  loc.stats[[k]] <- capture.recapture.stats(useful.site[k],
                                            dat.2.Cm %>% filter(Macro_site == useful.site[k]),
                                            save.fig = F)
}

```

In each loc.stats list item, a figure of the number of recaptures can be found in loc.stats[[k]]$plots$n.recap. From this plot, I select the starting season for each location. 


```{r}
loc.stats[[1]]$plots$n.recap


```

For BKS, the number of captures were very low until the winter of 2015 (2015-01). So, I removed the first four seasons.

```{r}
dat.2.Cm %>% filter(Macro_site == "BKS") %>%
  filter(season %ni% c("2012-08", "2013-08", "2014-01", "2014-08")) -> dat.3.Cm.BKS

loc.stats.BKS <- capture.recapture.stats("BKS",
                                         dat.3.Cm.BKS,
                                         save.fig = F)

loc.stats.BKS$plots$n.recap
```


```{r}
loc.stats[[2]]$plots$n.recap

```

For CI_IES, the first season (2005-08) will be removed.

```{r}

dat.2.Cm %>% filter(Macro_site == "CI_IES") %>%
  filter(season %ni% c("2005-08")) -> dat.3.Cm.CI_IES

loc.stats.CI_IES <- capture.recapture.stats("CI_IES",
                                            dat.3.Cm.CI_IES,
                                            save.fig = F)

loc.stats.CI_IES$plots$n.recap


```


```{r}
loc.stats[[3]]$plots$n.recap


```

CL_BMA looks good. 

```{r}
loc.stats[[4]]$plots$n.recap


```

This site needs to be removed. Not possible to estimate abundance from these captures. 

```{r}
useful.site[useful.site != "CPU"] -> useful.site.1 
```


```{r}
loc.stats[[5]]$plots$n.recap


```

For GNO, it's difficult to see how many were caught in the first few years from 
this plot.

```{r}

dat.2.Cm %>% filter(Macro_site == "GNO") %>%
  filter(DATE < as.Date("2004-01-01")) -> temp.GNO

loc.stats.temp.GNO <- capture.recapture.stats("GNO",
                                            temp.GNO,
                                            save.fig = F)

loc.stats.temp.GNO$plots$n.recap

```

This may be okay... 



```{r}
loc.stats[[6]]$plots$n.recap


```

For MUL, I remove the first two seasons.

```{r}

dat.2.Cm %>% filter(Macro_site == "MUL") %>%
  filter(season %ni% c("2006-01", "2008-01")) -> dat.3.Cm.MUL

loc.stats.MUL <- capture.recapture.stats("MUL",
                                         dat.3.Cm.MUL,
                                         save.fig = F)

loc.stats.MUL$plots$n.recap


```


```{r}
loc.stats[[7]]$plots$n.recap


```


PAO looks good. So, starting seasons for these sites are:

BKS = 2015-01
CI_IES = 2006-01
CL_BMA = 2001-08
GNO = 2001-08
MUL = 2008-08
PAO = 2001-08




