---
title: "Capture recapture analyses for Agnese Mancini"
output: html_notebook
---

This document describes data analysis of Agnese Mancini's data of hawksbill turtle capture recapture events at Isla Espiritu Santo (IES) using Mark. 


Initialize the workspace
```{r}
rm(list=ls())
library(jagsUI)
library(tidyverse)
library(lubridate)
library(reshape)
library(bayesplot)
library(ggridges)
library(RMark)
library(R2ucare)

# clean up the Mark output files before running the new ones.
cleanup(ask = F)
source("Mancini_functions.R")

do_analysis <- function(dp, ddl)
{
  # create formulas for Phi
  # tsm is time-since-marking; check for transient effects
  Phi.dot <-  list(formula = ~ 1)  
  Phi.CCL <- list(formula= ~ min_CCL)   # only 3 missing data for Ei so let's include this
  #Phi.t <- list(formula = ~ time)             # we never have this model worked for turtles... 
  #Phi.season <- list(formula = ~ sum_win)      # this also is unlikely... 
  #Phi.transience <- list(formula = ~ Transient)
  Phi.tsm <- list(formula = ~ tsm)
  
  #create formulas for p
  p.dot <- list(formula = ~ 1)
  p.t <- list(formula = ~ time)
  p.tsm <- list(formula = ~ tsm)
  p.transience <- list(formula = ~ Transient)
  #p.tsm.transience <- list(formula = ~ tsm + Transient)
  #p.t.transience <- list(formula = ~ time + Transient)
  p.effort <- list(formula = ~ effort)
  p.season <- list(formula = ~ sum_win)
  p.tsm.season <- list(formula = ~ tsm + sum_win)
  p.tsm.effort <- list(formula = ~ tsm + effort)
  
  # create all combinations 
  cml <- create.model.list("CJS")
  
  # run all all models and return as a list with class marklist
  results <- mark.wrapper(cml,
                          data=dp,
                          ddl=ddl,
                          output=FALSE,
                          silent=TRUE)
  return(results)
}

```

Bring in the data file and get ready for CJS.  There are few captures at the beginning and some no capture seasons also.  I wonder if this may be an issue for running CJS... 

```{r warning=F}
dat.1.Ei <- get.data.Ei("data/GTC_ICAPO_IES.csv")

CJS.data <- dat2CJS(dat.1.Ei, save.file = FALSE)

CJS.data$data %>% rownames_to_column(var = "ID") -> CH.1 #data.CJS

# using R2ucare::group_data to combine CHs - with individual covariates,
# this is not useful
CH.Ucare <- R2ucare::group_data(CJS.data$data, rep(1, nrow(CJS.data$data)))

# Use CCL as a covariate 
dat.1.Ei %>% select(ID, CCL) %>% group_by(ID) %>%
  summarise(min_CCL = min(CCL, na.rm = T)) %>%
  filter(!is.infinite(min_CCL))  -> data.CCL

# This gets you just 0s and 1s
data.CCL %>% left_join(CH.1, by = "ID") %>%
  select(-c("ID", "min_CCL")) -> CH.2

data.CCL  %>% left_join(CH.1, by = "ID") %>%
  select(min_CCL) -> cov.CCL

# here I define transients to be those that were caught just once
# but that eliminates the possibility of dying after the first
# capture - how do I deal with this?
n.cap <- rowSums(CJS.data$data)
transient.vec <- rep(1, times = length(n.cap))
transient.vec[n.cap > 1] <- 2

# need to count how many capture events occured per season
dat.1.Ei %>% select(season, "DATE") %>% 
  group_by(season) %>% #-> tmp3
  summarise(effort = n_distinct(DATE)) -> effort.season

# capture history
tmp <- apply(as.matrix(CH.2), 
             MARGIN = 2,
             FUN = paste0)

CH <- unite(data.frame(tmp),
           col = "ch",
           sep = "")

# with individual covariates, frequency is 1.  
#CH$freq <- CH.2$effY

# tmp <- apply(as.matrix(CJS.data$data), 
#              MARGIN = 2, 
#              FUN = paste0)

# CH <- unite(data.frame(tmp), 
#             col = "ch", 
#             sep = "")

#CH$min_weight <- cov.weight
CH$min_CCL <- cov.CCL
CH$Transient <- transient.vec


# capture dates and difference in years
cap.dates <- as.Date(paste0(colnames(CJS.data$data), "-01"))
# delta.dates <- signif(as.numeric(as.Date(cap.dates[2:length(cap.dates)]) -
#                                    as.Date(cap.dates[1:(length(cap.dates)-1)]))/365, 1)

delta.dates <- signif(as.numeric(cap.dates[2:length(cap.dates)] - 
                                   cap.dates[1:(length(cap.dates)-1)])/365, 1)

# a possibility of survival changing between summer/winter?
tmp <- strsplit(colnames(CJS.data$data), split = "-")
tmp2 <- lapply(tmp, FUN = function(x) x[2])
tmp3 <- unlist(tmp2)

# dp <- process.data(CH,
#                    model = "CJS",
#                    time.intervals = delta.dates,
#                    groups = "Transient",
#                    begin.time = min(year(cap.dates)))


# EFFORT AND TRANSIENT ARE NOT IN THE DATA.... FIX IT HERE 1/30/2020
dp <- process.data(CH,
                   model = "CJS",
                   time.intervals = delta.dates,
                   begin.time = min(year(cap.dates)))

ddl <- make.design.data(dp)

# p is indexed from 2010.6
# Phi is indexed from 2010

# effort affects the capture probability (p)
effort.season <- effort.season[1:length(levels(ddl$p$time)),]
effort.season$time <- as.factor(min(year(cap.dates)) + cumsum(delta.dates))

ddl$p <- merge_design.covariates(ddl$p, df = effort.season)

# summer/winter affects capture probability (maybe)
sum_win <- data.frame(sum_win = unlist(ifelse(tmp3 == "08", "summer", "winter")) [1:length(levels(ddl$p$time))],
                      time = as.factor(min(year(cap.dates)) + cumsum(delta.dates)))

ddl$p <- merge_design.covariates(ddl$p, df = sum_win)

# summer/winter affects survival (maybe)
sum_win <- data.frame(sum_win = unlist(ifelse(tmp3 == "08", "summer", "winter")) [1:length(levels(ddl$p$time))],
                      time = as.factor(min(year(cap.dates)) + 
                                         cumsum(c(0, delta.dates[1:(length(delta.dates)-1)]))))

ddl$Phi <- merge_design.covariates(ddl$Phi, df = sum_win)

# add time-since-marking (TSM models) - not sure if I'm doing this right... 
# When age = 0, it is the first time turtles were caught group(1)
# all others are group(2), which are recaptures. 
ddl$Phi$tsm <- 1
ddl$Phi$tsm[ddl$Phi$age == 0] <- 2

ddl$p$tsm <- 1
ddl$p$tsm[ddl$p$age == 0] <- 2

ei.results <- do_analysis(dp = dp, ddl = ddl)

```

Compare results using AICc:

```{r}
model.table(ei.results)
```

According to AICc, the best one is Phi(tsm)p(Transient). But c-hat > 11!  But Fletcher's chat is ~1.0.  Obviously, this isn't a great model... What to do? 

Look at the estimates:
```{r}
ei.results$Phi.tsm.p.transience$results$real
```

Export to Mark to see Release GOF.
```{r}
export.MARK(dp, 
            project.name = "Mancini_Ei_IES", 
            model = ei.results$Phi.tsm.p.transience, 
            replace = T, 
            chat = 11)
```

Because of the individual covariate, median c-hat analysis can't be run. 

R2ucare to do GOF.  There is no GOF for individual covariate models so we need to remove them before checking GOF (from user guide for R2ucare).

```{r}

#CH.ucare <- select(CH, -c("min_weight", "Transient"))
# test away! 
test3sr_Ei <- test3sr(as.matrix(CH.Ucare[, 1:(ncol(CH.Ucare)-1)]), 
                      freq = CH.Ucare$effY) #[, 1:(ncol(CH.2)-1)]), CH.2$effY)
test3sm_Ei <- test3sm(as.matrix(CH.Ucare[, 1:(ncol(CH.Ucare)-1)]), 
                      freq = CH.Ucare$effY)

test2ct_Ei <- test2ct(as.matrix(CH.Ucare[, 1:(ncol(CH.Ucare)-1)]), 
                      freq = CH.Ucare$effY)
test2cl_Ei <- test2cl(as.matrix(CH.Ucare[, 1:(ncol(CH.Ucare)-1)]), 
                      freq = CH.Ucare$effY)

# look at the overall results
test_all_Ei <- overall_CJS(as.matrix(CH.Ucare[, 1:(ncol(CH.Ucare)-1)]), 
                      freq = CH.Ucare$effY)

# using Justin's equations from Howick's data:
stat_new <- overall_CJS(as.matrix(CJS.data$data),
                        rep(1, nrow(CJS.data$data)))$chi2 -
  (test3sr(as.matrix(CJS.data$data),
           rep(1, nrow(CJS.data$data)))$test3sr[[1]])
df_new <- overall_CJS(as.matrix(CJS.data$data),
                      rep(1, nrow(CJS.data$data)))$degree_of_freedom -
  (test3sr(as.matrix(CJS.data$data), rep(1, nrow(CJS.data$data)))$test3sr[[2]])

1-pchisq(stat_new, df_new)

```

So... this is supposed to be okay?  Need to look into more on this stuff... 


Look at one at a time:
```{r}
test3sr.details <- test3sr_Ei$details

filter(test3sr.details, p_val < 0.05)
```

What's up with all these zeros? 

TEST3.SM, on the other hand, looks at individuals who were seen again.  "Among these individuals seen again, when they were seen again does not depend on whether or not they were seen for the first time at occasion (i). So, let's take a look at that result.

```{r}
test3sm.details <- test3sm_Ei$details
filter(test3sm.details, p_val < 0.05)
```

All zeros!!  What's going on here?




Moving on to TEST2... TEST2.CT tests the hypothesis that "there is no difference in the probability of being recaptured at t+1 between those captured and not captured at occasion t, conditional on presence at both occasions.  

```{r}
test2ct_Ei.details <- test2ct_Ei$details
filter(test2ct_Ei.details, p_val < 0.05)
```



TEST2.CL tests if there is no difference in the expected time of next capture between the individuals captured and not captured at occasion t conditional on presence at both occasions t adn t + 2.  

```{r}
test2cl_Ei.details <- test2cl_Ei$details
filter(test2cl_Ei.details, p_val < 0.05)
```



```{r}
test_all_Ei
```

These tests failed at some occasions. Does this mean the entire model is not suited to the data, or there are some occasions that failed the test but the CJS model isn't a complete failure because given the null-hypothesis approach, some tests are bound to fail when there are many occasions?  I see a good sign that many occasions passed these tests... The main concern right now is the estimated c-hat value of > 11.  

How do we deal with that? 

It's possible that using the other criteria, such as median c-hat and other GOF measures, we may be able to get away with using the CJS model. Assuming, that can be done... look at abundance. Do I have to separate residents from transients (only seen just once)?

The estimated proportion of residents can be computed from the two survival rates:
```{r}
p.residents <- ei.results$Phi.tsm.p.t$results$real$estimate[1]/ei.results$Phi.tsm.p.t$results$real$estimate[2]
  
phats <- ei.results$Phi.tsm.p.t$results$real$estimate[3:38]
n.caught <- colSums(CJS.data$data)
Nhats <- (n.caught[2:length(n.caught)]/phats) * p.residents
```

These seem so high! Are there that many green turtles in the area? 


