---
title: "von Bertalanffy analysis summary"
output: html_notebook
---


```{r}
rm(list = ls())
library(ggplot2)
library(tidyverse)
library(bayesplot)
library(rjags)
library(loo)

source("Mancini_functions.R")
```


This document summarizes results of somatic growth analysis for Agnese Mancini's capture-recapture data of marine turtles. Two species were analyzed: greens and hawksbills.

Retrieve data first:

```{r}
dat.1.Cm <- get.data.Cm("data/GTC_Cm Data_updated_2020-04-28_TE_v2.csv")

#dat.1 <- get.data("data/GTC_20190725_Tomo_v2.csv")

# Lo doesn't have enough (>2) recaptures. 
#spp <- c("Cm", "Ei") #unique(dat.1$species)

  
```


First greens:
```{r}
jags.data <- vonBert.jags.data(dat.1.Cm, "Cm")

jm.RLinf_Rk.Cm <- readRDS("RData/Rlinf_Rk_Cm_Aug2020.rds")
jm.RLinf_Fk.Cm <- readRDS("RData/Rlinf_Fk_Cm_Aug2020.rds")
jm.FLinf_Rk.Cm <- readRDS("RData/Flinf_Rk_Cm_Aug2020.rds")
jm.FLinf_Fk.Cm <- readRDS("RData/Flinf_Fk_Cm_Aug2020.rds")

# DIC.df <- data.frame(model.names = c("RLinf_Rk", "RLinf_Fk", "FLinf_Rk", "FLinf_Fk"),
#                      DIC = c(jm.RLinf_Rk.Cm$DIC, jm.RLinf_Fk.Cm$DIC, jm.FLinf_Rk.Cm$DIC, jm.FLinf_Fk.Cm$DIC)) %>%
#   arrange(by = DIC)

LOOIC.RLinf_Rk.Cm <- readRDS("RData/LOO_Rlinf_Rk_Cm_Aug2020.rds")
LOOIC.RLinf_Fk.Cm <- readRDS("RData/LOO_Rlinf_Fk_Cm_Aug2020.rds")
LOOIC.FLinf_Rk.Cm <- readRDS("RData/LOO_Flinf_Rk_Cm_Aug2020.rds")
LOOIC.FLinf_Fk.Cm <- readRDS("RData/LOO_Flinf_Fk_Cm_Aug2020.rds")

LOOIC.DIC.Cm <- data.frame(model = c("RLinf_Rk", "RLinf_Fk", "FLinf_Rk", "FLinf_Fk"),
                           LOOIC = c(LOOIC.RLinf_Rk.Cm$loo.out$estimates["looic", "Estimate"],
                                     LOOIC.RLinf_Fk.Cm$loo.out$estimates["looic", "Estimate"],
                                     LOOIC.FLinf_Rk.Cm$loo.out$estimates["looic", "Estimate"],
                                     LOOIC.FLinf_Fk.Cm$loo.out$estimates["looic", "Estimate"]),
                           SE = c(LOOIC.RLinf_Rk.Cm$loo.out$estimates["looic", "SE"],
                                  LOOIC.RLinf_Fk.Cm$loo.out$estimates["looic", "SE"],
                                  LOOIC.FLinf_Rk.Cm$loo.out$estimates["looic", "SE"],
                                  LOOIC.FLinf_Fk.Cm$loo.out$estimates["looic", "SE"]),
                           DIC = c(jm.RLinf_Rk.Cm$DIC, 
                                   jm.RLinf_Fk.Cm$DIC, 
                                   jm.FLinf_Rk.Cm$DIC, 
                                   jm.FLinf_Fk.Cm$DIC),
                           max.Rhat = c(max(jm.RLinf_Rk.Cm$summary %>% as.data.frame() %>% select(Rhat)),
                                        max(jm.RLinf_Fk.Cm$summary %>% as.data.frame() %>% select(Rhat)),
                                        max(jm.FLinf_Rk.Cm$summary %>% as.data.frame() %>% select(Rhat)),
                                        max(jm.FLinf_Fk.Cm$summary %>% as.data.frame() %>% select(Rhat)))) %>%
  arrange(by = LOOIC)

LOOIC.DIC.Cm
```


According to LOOIC, random Linf and fixed k parameter model was considered best for green turtles. Look at the Pareto-k statistic for the goodness-of-fit.

```{r}
pareto_k_table(LOOIC.RLinf_Fk.Cm$loo.out)
```

Some are not very good but 95% of data were ok or good. So, moving on with the model. 

Take a look at some posteriors to see if convergence was okay.  According to the Rhat statistics, they were okay.

```{r}
mcmc_trace(jm.RLinf_Fk.Cm$samples, c("LinfMu", "k", "Shape", "rate"))
```

Not great but they look okay. 


```{r}
mcmc_dens(jm.RLinf_Fk.Cm$samples, c("LinfMu", "k", "Shape", "rate"))
```

Next, I plot the fitted von Bertalanffy function. Need to use the joint posterior, not parameter-wise credible intervals.
```{r}
k.samples <- extract.samples("k", jm.RLinf_Fk.Cm$samples)

time.vec <- 0:59

unique.ID <- jags.data$ID

time.mat <- L_Exp.med <- L_Exp.lo <- L_Exp.hi <- matrix(ncol = length(unique.ID), nrow = length(time.vec))

t <- i <- 1
for (i in 1:length(unique.ID)){
  A.samples <- extract.samples(paste0("A[", i, "]"), jm.RLinf_Fk.Cm$samples)
  Linf.samples <- extract.samples(paste0("Linf[", i, "]"), jm.RLinf_Fk.Cm$samples)
  for (t in 1:length(time.vec)){
    L_Exp <-  Linf.samples * (1.0 - exp(-k.samples * (A.samples + time.vec[t])))
    L_Exp.med[t,i] <- quantile(L_Exp, 0.5)
    L_Exp.lo[t,i] <- quantile(L_Exp, 0.025)
    L_Exp.hi[t,i] <- quantile(L_Exp, 0.975)
    
    time.mat[t,i] <- quantile(A.samples + time.vec[t], 0.5)
  }

}

colnames(L_Exp.med) <- unique.ID
L_Exp_long <- melt(L_Exp.med)
colnames(L_Exp_long) <- c("time1", "ID", "CCL")

colnames(L_Exp.lo) <- unique.ID
L_Exp.lo_long <- melt(L_Exp.lo)
colnames(L_Exp.lo_long) <- c("time1", "ID", "CCL")

colnames(L_Exp.hi) <- unique.ID
L_Exp.hi_long <- melt(L_Exp.hi)
colnames(L_Exp.hi_long) <- c("time1", "ID", "CCL")

colnames(time.mat) <- unique.ID
time_long <- melt(time.mat)
colnames(time_long) <- c("time1", "ID", "Years")

L_Exp_long$Years <- time_long$Years
L_Exp.lo_long$Years <- time_long$Years
L_Exp.hi_long$Years <- time_long$Years

ggplot() + 
  geom_path(data = L_Exp_long,
            aes(x = Years, y = CCL, color = ID)) +
  geom_path(data = L_Exp.lo_long,
            aes(x = Years, y = CCL, color = ID),
            linetype = 2) +
  geom_path(data = L_Exp.hi_long,
            aes(x = Years, y = CCL, color = ID),
            linetype = 2) +
  
  theme(legend.position = "none")
```

