---
title: "von Bertalanffy analysis summary"
output: html_notebook
---


```{r}
rm(list = ls())
library(ggplot2)
library(tidyverse)
library(bayesplot)
library(rjags)

source("Mancini_functions.R")
```


This document summarizes results of somatic growth analysis for Agnese Mancini's capture-recapture data of marine turtles. Two species were analyzed: greens and hawksbills.

Retrieve data first:

```{r}
dat.1 <- get.data("data/GTC_20190725_Tomo_v2.csv")

# Lo doesn't have enough (>2) recaptures. 
spp <- c("Cm", "Ei") #unique(dat.1$species)

  
```


First greens:
```{r}
jags.data <- vonBert.jags.data(dat.1, "Cm")

jm.RLinf_Rk.Cm <- readRDS("RData/Rlinf_Rk_Cm.rds")
jm.RLinf_Fk.Cm <- readRDS("RData/Rlinf_Fk_Cm.rds")
jm.FLinf_Rk.Cm <- readRDS("RData/Flinf_Rk_Cm.rds")
jm.FLinf_Fk.Cm <- readRDS("RData/Flinf_Fk_Cm.rds")

DICs <- c(jm.RLinf_Rk.Cm$DIC, jm.RLinf_Fk.Cm$DIC, jm.FLinf_Rk.Cm$DIC, jm.FLinf_Fk.Cm$DIC)

```

Although, the usefulness of DIC has been questioned, fixed Linf and fixed k parameter model was considered best for green turtles.  Take a look at some posteriors to see if convergence was okay.  According to the Rhat statistics, they were okay.

```{r}
mcmc_trace(jm.FLinf_Fk.Cm$samples, c("Linf", "Shape", "rate"))
```

They look okay. 


```{r}
mcmc_dens(jm.FLinf_Fk.Cm$samples, c("Linf", "Shape", "rate", "k"))
```


Next, I plot the fitted von Bertalanffy function. 
```{r}
Linf <- jm.FLinf_Fk.Cm$q50$Linf
k <- jm.FLinf_Fk.Cm$q50$k
A <- jm.FLinf_Fk.Cm$q50$A

Linf.lo <- jm.FLinf_Fk.Cm$q2.5$Linf
k.lo <- jm.FLinf_Fk.Cm$q2.5$k
A.lo <- jm.FLinf_Fk.Cm$q2.5$A

Linf.hi <- jm.FLinf_Fk.Cm$q97.5$Linf
k.hi <- jm.FLinf_Fk.Cm$q97.5$k
A.hi <- jm.FLinf_Fk.Cm$q97.5$A

time.vec <- 0:59

unique.ID <- jags.data$ID

time.mat <- L_Exp <- L_Exp.lo <- L_Exp.hi <- matrix(ncol = length(unique.ID), nrow = length(time.vec))

i <- 1
for (i in 1:length(unique.ID)){
  L_Exp[,i] <-  Linf * (1.0 - exp(-k * (A[i] + time.vec)))
  L_Exp.lo[,i] <-  Linf.lo * (1.0 - exp(-k * (A[i] + time.vec)))
  L_Exp.hi[,i] <-  Linf.hi * (1.0 - exp(-k * (A[i] + time.vec)))
  
  time.mat[,i] <- A[i] + time.vec
}

colnames(L_Exp) <- unique.ID
L_Exp_long <- melt(L_Exp)
colnames(L_Exp_long) <- c("time1", "ID", "CCL")

colnames(L_Exp.lo) <- unique.ID
L_Exp.lo_long <- melt(L_Exp.lo)
colnames(L_Exp.lo_long) <- c("time1", "ID", "CCL")

colnames(L_Exp.hi) <- unique.ID
L_Exp.hi_long <- melt(L_Exp.hi)
colnames(L_Exp.hi_long) <- c("time1", "ID", "CCL")

colnames(time.mat) <- unique.ID
time_long <- melt(time.mat)
colnames(time_long) <- c("time1", "ID", "Years")

L_Exp_long$Years <- time_long$Years
L_Exp.lo_long$Years <- time_long$Years
L_Exp.hi_long$Years <- time_long$Years

ggplot() + 
  geom_path(data = L_Exp_long,
            aes(x = Years, y = CCL, color = ID)) +
  geom_path(data = L_Exp.lo_long,
            aes(x = Years, y = CCL, color = ID),
            linetype = 2) +
  geom_path(data = L_Exp.hi_long,
            aes(x = Years, y = CCL, color = ID),
            linetype = 2) +
  
  theme(legend.position = "none")
```

